On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages

### 2020-10-03 paper/paper.tex
Convert to PADL format (LNCS)

%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,anonymous,review]{acmart}
\usepackage[switch]{lineno}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{tcolorbox}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{stmaryrd}
\usepackage{amsmath}

@@

\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{XX.XXXX/XXXXXXX}

@@

\acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
  ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
\acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
Conference on Generative Programming: Concepts and Experiences (GPCE '20)
  November 15 -- 20, 2020, Chicago, Illinois, United States}
\acmPrice{15.00}
\acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}

\citestyle{acmauthoryear}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[On Adding Pattern Matching to Haskell-based EDSLs]{On Adding Pattern Matching to Haskell-based
  Deeply Embedded Domain Specific Languages}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{David Young}
\email{d063y800@ku.edu}

% \authornote{Both authors contributed equally to this research.}
% \email{trovato@corporation.com}
% \orcid{1234-5678-9012}

\author{Mark Grebe}
\email{grebe@ucmo.edu}

\author{Andrew Gill}
\email{andygill@ku.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
\renewcommand{\shortauthors}{Young, Grebe and Gill}

@@

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011041.10011046</concept_id>
       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011041.10011047</concept_id>
       <concept_desc>Software and its engineering~Source code generation</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011050.10011017</concept_id>
       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
\ccsdesc[300]{Software and its engineering~Source code generation}
\ccsdesc[500]{Software and its engineering~Domain specific languages}

@@

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

@@

\newpage
Haskell EDSLs that support powerful language features like pattern matching
allow these advantages to be brought to problem domains where it is currently
either difficult or impossible to use a language that has these benefits.

@@

\newpage
\begin{lstlisting}
data E t where
  ...
  SumMatchExp ::
    (ERep a, ERep b, ERepTy b ~ b)
      => E (ProdMatch a r)
         -> E (SumMatch b r)
         -> E (SumMatch (Either a b) r)

@@

  Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
\end{todo}

@@

These bindings are grouped together in the \ttt{Env} type:
\newpage
\begin{lstlisting}
newtype Env = Env [EnvMapping]

@@

\newpage
\section{C Backend}
\label{sec:CBackend}

@@

To use this, for example, with the previously mentioned \ttt{isEven} implementation we
can write:
\newpage
\begin{lstlisting}
isEvenE :: E (Int -> Bool)
isEvenE =
  externalize (
    case x == 0 of
      True -> True
      False ->
        case x == 1 of
          True -> False
          False -> isEvenE (x - 2))

@@

\clearpage
\begin{figure*}
  \centering 
  \resizebox{2.0\columnwidth}{20cm}{\includegraphics{Diagrams/Flowchart.pdf}}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{figure*}

\clearpage

@@

  &D\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}
      \rewrites I(D(\ttt{f x = externalize (case s of \{ ... \})}))\\

@@

\captionsetup{justification=centering}
\caption{Rewrite rules}
\label{fig:Rewrites}
\end{figure*}

@@

\clearpage

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

>>>>

%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\RequirePackage{amsmath,amssymb}
% \RequirePackage{stmaryrd}
\documentclass[runningheads, a4paper]{llncs}
% \usepackage{biblatex}
% \usepackage[switch]{lineno}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{tcolorbox}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{stmaryrd}
% \usepackage{amsmath}

@@

% \setcopyright{acmcopyright}
% \copyrightyear{2020}
% \acmYear{2020}
% \acmDOI{XX.XXXX/XXXXXXX}

@@

% \acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
%   ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
% \acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
% Conference on Generative Programming: Concepts and Experiences (GPCE '20)
%   November 15 -- 20, 2020, Chicago, Illinois, United States}
% \acmPrice{15.00}
% \acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}

% \citestyle{acmauthoryear}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{On Adding Pattern Matching to Haskell-based
  Deeply Embedded Domain Specific Languages}

\titlerunning{On Adding Pattern Matching to Haskell-based EDSLs}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andrew Gill\inst{1}}
\institute{University of Kansas \and University of Central Missouri}

\maketitle
% \authornote{Both authors contributed equally to this research.}
% \email{trovato@corporation.com}
% \orcid{1234-5678-9012}

% \author{Mark Grebe}
% \email{grebe@ucmo.edu}

% \author{Andrew Gill}
% \email{andygill@ku.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

% \renewcommand{\shortauthors}{Young, Grebe and Gill}

@@

% \begin{CCSXML}
% <ccs2012>
%    <concept>
%        <concept_id>10011007.10011006.10011041.10011046</concept_id>
%        <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
%        <concept_significance>300</concept_significance>
%        </concept>
%    <concept>
%        <concept_id>10011007.10011006.10011041.10011047</concept_id>
%        <concept_desc>Software and its engineering~Source code generation</concept_desc>
%        <concept_significance>300</concept_significance>
%        </concept>
%    <concept>
%        <concept_id>10011007.10011006.10011050.10011017</concept_id>
%        <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
%        <concept_significance>500</concept_significance>
%        </concept>
%  </ccs2012>
% \end{CCSXML}

% \ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
% \ccsdesc[300]{Software and its engineering~Source code generation}
% \ccsdesc[500]{Software and its engineering~Domain specific languages}

@@

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

@@

% \newpage
Haskell EDSLs that support powerful language features like pattern matching
allow these advantages to be brought to problem domains where it is currently
either difficult or impossible to use a language that has these benefits.

@@

% \newpage
\begin{lstlisting}
data E t where
  ...
  SumMatchExp ::
    (ERep a, ERep b, ERepTy b ~ b)
      => E (ProdMatch a r)
         -> E (SumMatch b r)
         -> E (SumMatch (Either a b) r)

@@

  % Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
\end{todo}

@@

These bindings are grouped together in the \ttt{Env} type:
% \newpage
\begin{lstlisting}
newtype Env = Env [EnvMapping]

@@

% \newpage
\section{C Backend}
\label{sec:CBackend}

@@

To use this, for example, with the previously mentioned \ttt{isEven} implementation we
can write:
% \newpage
\begin{lstlisting}
isEvenE :: E (Int -> Bool)
isEvenE =
  externalize (
    case x == 0 of
      True -> True
      False ->
        case x == 1 of
          True -> False
          False -> isEvenE (x - 2))

@@

% \clearpage
\begin{figure*}
  \centering 
  \resizebox{0.8\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
  % \includegraphics{Diagrams/Flowchart.pdf}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{figure*}

@@

  &D\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
      &\quad\rewrites I(D(\ttt{f x = externalize (case s of \{ ... \})}))\\

@@

% \captionsetup{justification=centering}
\caption{Rewrite rules}
\label{fig:Rewrites}
\end{figure*}

@@

@@

\bibliographystyle{splncs04}
\bibliography{paper}
